A Transdiciplinary approach to developing a spectral numerical coastal-flow

### 2019-08-29 literature_review.Rmd
house cleaning

---
title: 'A Transdiciplinary approach to developing a spectral numerical coastal-flow
  model based on passive macroalgal rafting phenenomea to shed light on ambient and
  stochastic dispersal and connectivity of nearshore and offshore marine communities
  hydrodynamic modelling in and around the coast of South Africa: Combining ecology and 3D hyperspectral'
author: "Ross Coppin"
output:
  word_document:
    fig_caption: yes
    toc: yes
---

```{r include=FALSE}
library(png)
library(grid)
```

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling eﬀect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow eﬀects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and * Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure ???). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure ??). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a wave into account (the front, see figure ??) which is relevant to characterising breaking waves. It should be noted that surface elevation can be negative while a wave profile cannot.

```{r surface elevation figures, echo=FALSE, warning=FALSE, fig.cap="The definition of a wave in a time record of the surface elevation with downward zero-crossings (upper panel) and upward zero-crossings (lower panel)."}

>>>>

---
title: 'A Transdiciplinary approach to developing a spectral numerical coastal-flow
  model based on passive macroalgal rafting phenenomea to shed light on ambient and
  stochastic dispersal and connectivity of nearshore and offshore marine communities
  and 3D hyperspectral hydrodynamic modelling in and around the coast of South Africa'
author: "Ross Coppin"
output:
  pdf_document: 
  word_document:
    fig_caption: yes
    reference_docx: frontiers_template.docx
    toc: yes
---

```{r include=FALSE}
library(png)
library(grid)
library(rticles)
```

@@

Other than temperature, wave exposure is also recognised as an important driver of the marine environment, and macroalgae are not exception. Wave exposure has been shown to play a role in determining distribution, abundance, diversity, composition, growth (Cousens 1982) and productivity (Pedersen and Nejrup 2012) of macroalgae communities. For example, the width, vertical zonation and diversity of algal communities often change predictably along gradients of wave exposure. Wave exposure may also drive macroalgae communities indirectly through the alteration in effect of another environmental driver. For instance, increasing degrees of exposure may positively influence the amount of area available to trap light on macroalgal fronds, as well as increasing nutrient uptake due to increased turbulence in the boundary layer around the frond (???). The most important direct effect of wave exposure on macroalgal communities is through mechanical dislodgement, which ultimately leads to expiration. Wave exposure is a complex abiotic variable which varies spatially and temporarily in the marine environment. Furthermore, the degree to which a macroalgae community is exposed, is dependent on local site characteristics, such as bathymetry and local wind patterns. Despite this fact, macroalgae have been able to persist in often harsh and variable wave environments. Macroalgae are sessile organisms and incapabable of migrating when local conditions become unsuitable. Therefore, macroalgae must adapt to the local wave climate in order to persist and survive, and achieve this through morphological adaptation. The morphology of macroalgae are not fixed genetic traits. A study by Koehl et al. (2008) showed that transplanted Nereocystis luetkeana plants from a wave sheltered site to a wave exposed site changed their morphology to flat blades and narrow laterals that are less prone to drag forces in 4-5 days. Advances in genetic techniques and taxonomy have revealed that species delineation based on morphology has been inaccurate, and organisms that were once considered two separate species are actually one species. For example, Moss (1948) investigated the anatomy, chemical composition of Fucus spiralis at three sites that varied in wave exposure (sheltered, medium exposure and exposed). The authors found that individuals in exposed sites showed less branching of thalli as well as variation physiological components, such as organic nitrogen, mannitol, laminarian and alginic acid concentrations. The authors also noted a ‘crumpling effect’ displayed by individuals from exposed sites and inferred that this strategy may reduce overall drag. Other studies show that macroalgae in wave exposed environments have morphologies that reduce overall drag, increase strength of attachment or increase flexibility. There is also evidence that morphological adaptation is driven by currents, and in fact may be driving hydrological performance of macroalgae. Duggins et al. (2003) examined the direct and indirect flow effects on population dynamics, morphology and biomechanics of several understorey macroalgae species. These species included Costaria costata, Agarum fimbriatum, and Laminaria complanata and *Nereocystis luetkeana*. The results showed that in wave impacted sites (wave exposed) had higher rates of mortality, and no significance was found between survival of individuals and tidal or current velocity. The authors concluded that although tidal and current velocity did not play a significant role in determining kelp survival, it did play a role in morphological adaptation. The results from this study suggest that high current and tidal stresses are the main driver of kelp morphological adaptation. This in turn make those individuals more resilient to dislodgement to wave exposure. Wave exposure is stochastic in nature compared to tidal and ocean currents which are more regular in their frequency and magnitude. Therefore the regular forces of tidal and ocean currents may make kelp individuals more resilient to mechanical dislodgement over time.

@@

Waves are often thought of as an elevation of the sea surface from a specific point over a period of time but this is obvisiously not he case. This is known as *surface elevation* and is the instaneous elevation of the sea surface abouve a specific point in a time record (see figure \ref{fig:surface elevation figures}). Although surface elevation does not represent a wave, it can be used to create a wave profile. This is achieved by profiling the the surface elevation between two successive *downward zero-crossings* or *upwardward zero-crossings* (see figure \ref{fig:surface elevation figures}). Both zero-crossings are symmetrical and essentially the same statistically. However, in practice the downward zero-crossings are preferred as it takes steepness of a wave into account (the front, see figure \ref{fig:surface elevation figures}) which is relevant to characterising breaking waves. It should be noted that surface elevation can be negative while a wave profile cannot.

```{r surface elevation figures, echo=FALSE, warning=FALSE, fig.cap="\\label{fig:surface elevation figures}The definition of a wave in a time record of the surface elevation with downward zero-crossings (upper panel) and upward zero-crossings (lower panel)."}
